Untwisting the Tetrahelix (v0.6)

### 2017-02-24 doc/TetrahelixUntwist.tex
improved references

\title{Untwisting the Tetrahelix (v0.5)}
\author{Robert L. Read \texttt{read.robert@gmail.com} \and
  Robert Gatliff \texttt{robert@toubat.org}
}

@@

\begin{abstract}
  The Boerdijk--Coxeter helix (BC helix, or tetrahelix) is a face-to-face stack of regular tetrahedra.
  Considering the edges of these tetrahedra, the resulting structure is attractive and inherently rigid,
  and therefore interesting to architects, mechanical engineers,
  and robotocists.
  A formula that matches the visually apparent helices forming the outer rails of the tetrahelix is derived which
  is convenient for designers.
  This formula defines the vertices of tetrahelices of varying radius, pitch, and curvature, with the BC helix
  as a special case. 
  An addtional special case is of that of $0$ curvature or rail angle, which generates a \emph{tetrabeam}.
  A particular choice of paramaters defines a novel object, the  minimax member length-difference tetrabeam,
  the \emph{equitetrabeam}.
  A formula for computing the radius and z-axis travel that gives an optimal minimax tetrahelix for any rail angle
  between $0$ and $\rho_{bc}$ is given, thus defining a continuum of tetrahelices containing the equitetrabeam and
  BC helix.
  Utility and use for static and variable geometry truss/space frame design and robotics based on the are discussed.
\end{abstract}

@@

\begin{figure}[H] %float with two figures
  \label{closeup}
  \centering
     \includegraphics[width=0.95\textwidth]{figures/BCHelixCloseUp.png}
     \caption{BC Helix Close-up}
\end{figure}

\begin{figure}[H] %float with two figures
  \label{series}
  \centering
     \includegraphics[width=0.95\textwidth]{figures/TetrahelixSeries.png}
     \caption{A Series of Tetrahelices, BC Helix in Foreground.}
\end{figure}

BC helix does not rest on a plane in a simple way. It is convenient to
be able to ``untwist'' it and form a tetrahelix space frame that has a
flat planar surface. By making length changes in a certain way, we can
untwist a tetrahelix to form a \emph{tetrabeam} which has planar faces
and has, for example, an equilateral triangular profile.  In Figure
\ref{series}, the closest helix is the BC helix.

@@

Coxeter constructs the BC helix\cite{coxeter1985simplicial} as a repeated rotation and translation of the tetrahedra, showing the
rotation is:
\[
\theta = \arccos(-2/3) 
\]
and the translation:
\[
h_{bc} = 1/\sqrt{10}
\]

@@

From Robert Gray's site, repeating formula by H.S.M. Coxeter, for a counter-clockwise BC Helix:
\begin{equation}
  \label{graycoxeter}
V(n) =
\left [
  \begin{tabular}{c}
   $ r_{bc} \cos(n \theta) $\\
   $ r_{bc} \sin(n \theta) $\\
   $ n h_{bc}  $
  \end{tabular}
  \right ],
\text{where:}
  \begin{tabular}{c}
 $ r_{bc} = \frac{3\sqrt{3}}{10} $\\
 $ h_{bc} = 1/\sqrt{10} $ \\
 $ \theta = \arccos(-2/3) $ \\
  \end{tabular}      
\end{equation}
where $n$ represents each integer numbered node in succession on every colored rail.

The apparent rotation of tetrhedra of the same outer-edge color ($V_3$ relative from $V_0$ for example,
in \eqref{graycoxer}, is $3 \theta - 2\pi$.

@@

\begin{align*}
  1 &=  \sqrt{\frac{1}{9}  + r_{\rho}^2(-\frac{4 (\sin^2( \rho / 2))}{9} + \sin^2(\rho/3+ \frac{2\pi}{3})  + (1 - \cos(\rho/3 + \frac{2\pi}{3}))^2)} \\
  r_{opt} &= \frac{2}{\sqrt{\frac{9}{2} \cdot ( \sqrt{3} sin(\rho/3) + \cos(\rho/3)) + cos(\rho)+ 8 }} \numberthis  \label{eqrhoopt} \\
\end{align*}
(where denominator $  \neq 0$).

@@

The pitch of any tetrahelix 
where $\rho \neq 0$ is:
\[
p(\rho) = \frac{2 \pi  \cdot d}{\rho}
\]

As pitch increases smoothly with decreasing rail angle, it is easy to solve for rail angle
to acheive a given pitch, an ability valuable to engineers or architects who wish
to fit a tetrahelix to given dimension.

Perhaps surprisingly, the optimal untwisting is accomplished only by changing the length of
the two-hop member, leaving the one-hop memever and
rail length equivalenet within this continuum. However, it should be noted that
an engineer or architect may also use $H_{general}$ in the same way, and that minimax
length optimality is a mathematic notion only loosely related to the beauty and
utility of physical structures.

@@

Engineers has two fundamental mathamematical tools for constructing a tetrahelix.
They may use

@@

@@

Suppose $\overline{\rm C'D'}$ is the shortest edge. Increasing $G$ thereby improves our mimimum, up until the next shortest edge length, $\overline{\rm AC}$.
This may increase $\overline{\rm BC}$ and $\overline{\rm AD}$, but more slowly (???) than $\overline{\rm C'D'}$ is being increased, until  $\overline{\rm C'D'} =
\overline{\rm AC}$.
Note: This is sketchy, can we show that the derivative of our formula for CD  with respect to $G$ is really higher than $AD$? These derivatives will depend on $\rho$ slightly, which is exactly right!

>>>>

\title{Untwisting the Tetrahelix (v0.6)}
\author{Robert L. Read \texttt{read.robert@gmail.com} \and
  Robert Gatliff \texttt{robert@toubat.org}
}

@@

\begin{abstract}
  The Boerdijk--Coxeter helix (BC helix, or tetrahelix) is a
  face-to-face stack of regular tetrahedra.  Considering the edges of
  these tetrahedra, the resulting structure is attractive and
  inherently rigid, and therefore interesting to architects,
  mechanical engineers, and robotocists.  A formula that matches the
  visually apparent helices forming the outer rails of the tetrahelix
  is derived which is convenient for designers.  This formula defines
  the vertices of tetrahelices of varying radius, pitch, and
  curvature, with the BC helix as a special case.  An addtional
  special case is of that of $0$ curvature or rail angle, which
  generates a \emph{tetrabeam}.  A particular choice of paramaters
  defines a novel object, the minimax member length-difference
  tetrabeam, the \emph{equitetrabeam}.  A formula for computing the
  radius and z-axis travel that gives an optimal minimax tetrahelix
  for any rail angle between $0$ and $\rho_{bc}$ is given, thus
  defining a continuum of tetrahelices containing the equitetrabeam
  and BC helix. Numerically finding $\rho$ from the equation for
  pitch allows optimal tetrahelices of any pitch to be designed.
  An interactive tool for such design and experimentation is provided.
  Utility and use for static and variable geometry
  truss/space frame design and robotics based on the are discussed.
\end{abstract}

@@

\begin{figure}[H] %float with two figures
  \label{fig:closeup}
  \centering
     \includegraphics[width=0.95\textwidth]{figures/BCHelixCloseUp.png}
     \caption{BC Helix Close-up}
\end{figure}

BC helix does not rest on a plane in a simple way. It is convenient to
be able to ``untwist'' it and form a tetrahelix space frame that has a
flat planar surface. By making length changes in a certain way, we can
untwist a tetrahelix to form a \emph{tetrabeam} which has planar faces
and has, for example, an equilateral triangular profile. This paper
develops the equations needed to untwist the tetrahelix. All math
deveoped here is available in Javascript and demonstrated by an interactive
design website\cite{readtetrahelix}, from which Figures \ref{fig:closeup} and
\ref{fig:series} are taken.

\begin{figure}[H] %float with two figures
  \label{fig:series}
  \centering
     \includegraphics[width=0.95\textwidth]{figures/TetrahelixSeries.png}
     \caption{A Series of Tetrahelices, BC Helix in Foreground.}
\end{figure}

%% Note: This is not rendering figure two for some reason.
In Figure
\ref{fig:series}, the closest helix is the BC helix.

@@

H.S.M Coxeter constructs the BC helix\cite{coxeter1985simplicial} as a repeated rotation and translation of the tetrahedra, showing the
rotation is:
\[
\theta = \arccos(-2/3) 
\]
and the translation:
\[
h_{bc} = 1/\sqrt{10}
\]

@@

From R.W. Gray's site\cite{graytetrahelix}, repeating formula by Coxeter\cite{coxeter1985simplicial}, for a counter-clockwise BC Helix:
\begin{equation}
  \label{graycoxeter}
V(n) =
\left [
  \begin{tabular}{c}
   $ r_{bc} \cos(n \theta) $\\
   $ r_{bc} \sin(n \theta) $\\
   $ n h_{bc}  $
  \end{tabular}
  \right ],
\text{where:}
  \begin{tabular}{c}
 $ r_{bc} = \frac{3\sqrt{3}}{10} $\\
 $ h_{bc} = 1/\sqrt{10} $ \\
 $ \theta = \arccos(-2/3) $ \\
  \end{tabular}      
\end{equation}
where $n$ represents each integer numbered node in succession on every colored rail.

The apparent rotation of tetrhedra of the same outer-edge color ($V_3$ relative from $V_0$ for example,
in \eqref{graycoxeter}, is $3 \theta - 2\pi$.

@@

\begin{align*}
  1 &=  \sqrt{\frac{1}{9}  + r_{opt}^2(-\frac{4 (\sin^2( \rho / 2))}{9} + \sin^2(\rho/3+ \frac{2\pi}{3})  + (1 - \cos(\rho/3 + \frac{2\pi}{3}))^2)} \\
  r_{opt} &= \frac{2}{\sqrt{\frac{9}{2} \cdot ( \sqrt{3} sin(\rho/3) + \cos(\rho/3)) + cos(\rho)+ 8 }} \numberthis  \label{eqrhoopt} \\
\end{align*}
(where denominator $  \neq 0$).

@@

The pitch of any tetrahelix 
where $\rho \neq 0$ is:
\begin{equation}
  \label{pitcheqn}
p(\rho) = \frac{2 \pi  \cdot d}{\rho}
\end{equation}
For a fixed $z$-axis travel $d$, this is trivial. 
$z$-axis travel as per \eqref{dopt} pitch increases monotonically and smoothly with decreasing $\rho$, so
Equation \eqref{pitcheqn} can be easily solved numerically.
For a pitch at least $ \frac{3  \sqrt{\frac{2}{5}}  \pi}{\rho_{bc}} \approx 9.64 $,
and using \eqref{dopt}, produces minimax optimal tetrahelices.

In this way a rail angle can be chosen for any desired pitch, yield
the optimum radius, one-hop, and two-hop lengths an engineer needs to
construct a physical structure.

Perhaps surprisingly, the optimal untwisting is accomplished only by
changing the length of the two-hop member, leaving the one-hop memever
and rail length equivalenet within this continuum. However, it should
be noted that an engineer or architect may also use $H_{general}$
directly and interactively, and that minimax length optimality is a
mathematic notion only loosely related to the beauty and utility of
physical structures.

@@

@@

The BC Helix is the end point of a continuum of tetrahelices, the
other end point being an uncurved tetrahelix with equilateral cross
section, constructed by changing the length of only those members
crossing the outside rails after hopping over the nearest
vertex. Under the condition of miminum maximum length difference of
all members in the system, all such tetrahelices have vertices evenly
spaced along the axis generated by a simple equation.  A mechanical
machine, such as a robot or a variable-geometry truss, that can change
the length of its members, can thus twist and untwist itself by
changing the length of the appropriate members to achieve any point in
the continuum optimally. With a numeric solution, a design may choose
a rotation angle and member lengths to obtain any desired pitch.

@@

Suppose $\overline{\rm C'D'}$ is the shortest edge. Increasing $G$
thereby improves our mimimum, up until the next shortest edge length,
$\overline{\rm AC}$.  This may increase $\overline{\rm BC}$ and
$\overline{\rm AD}$, but more slowly than $\overline{\rm C'D'}$
is being increased, until $\overline{\rm C'D'} = \overline{\rm AC}$.
Note: This is sketchy, can we show that the derivative of our formula
for CD with respect to $G$ is really higher than $AD$? These
derivatives will depend on $\rho$ slightly, which is exactly right!
